---
title: 'EDA: Airbnb listings in New York'
author: "Bhushan Borude"
date: "2022-08-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exploratory Data Analysis

Click [here](https://www.kaggle.com/datasets/arianazmoudeh/airbnbopendata) for the source of dataset.

```{r include=FALSE}
# This chunk will run the library code, but wont include it in the html file.
# Knitting happens in a fresh R session, so if you have not loaded your packages in a code chunk, you'll get errors.

# Load necessary packages
library("tidyverse")

# the next chunk will be shown in the html file but r will not run the code.
```


```{r, eval=FALSE}
# Load necessary packages
library("tidyverse")
```

```{r}
# Load the data
airbnb <- read.csv("Airbnb_Open_Data.csv")
```

```{r}
# Check the structure of data frame
str(airbnb)

# Get a summary of data frame
summary(airbnb)
```

### Data Preparation
* Remove unwanted columns
* Rename the columns
* Remove 'NA' from columns :lat, long, instant_booking, construction_year
* Remove entries where host verification status is empty
* Remove entries where borough is empty
* Get the mean value of rating
* Replace NA in rating with mean value
* Get rid of '$' and "," from price and cast as integer
* Replace 'brookln' in borough by 'Brooklyn'
* Add a column for price range
* Remove 'NA' from price and price_range

```{r, include=FALSE}

# Remove unwanted columns
airbnb_ny <- airbnb %>% 
  select(id, host.id, host_identity_verified, neighbourhood.group, neighbourhood, lat, long, instant_bookable, cancellation_policy, room.type, Construction.year, price, review.rate.number)
  
# Rename the columns
airbnb_ny <- rename(airbnb_ny, listing_id = id, 
                    host_id = host.id,
                    host_verification_status = host_identity_verified,
                    borough = neighbourhood.group,
                    instant_booking = instant_bookable,
                    room_type = room.type,
                    construction_year = Construction.year, 
                    rating = review.rate.number)

# Remove entries with 'NA' from columns :lat, long, instant_booking, construction_year
airbnb_ny <- airbnb_ny %>% 
  filter(!is.na(lat) & !is.na(long) & !is.na(instant_booking) & !is.na(construction_year))

# Remove entries where host verification status is empty
airbnb_ny <- airbnb_ny %>% 
  filter(!host_verification_status == "")

# Remove entries where borough is empty
airbnb_ny <- airbnb_ny %>% 
  filter(!borough == "")

# Get the mean value of rating
airbnb_ny %>% 
  filter(!is.na(rating)) %>% 
  summarize(mean_rating = mean(rating))

# Replace NA in rating with mean value
airbnb_ny$rating[is.na(airbnb_ny$rating)] <- 3

# Get rid of $ and "," from price and cast as integer
airbnb_ny$price <- substr(airbnb_ny$price, 2, 6)
airbnb_ny$price <- gsub(",", "", airbnb_ny$price)
airbnb_ny$price <- as.integer(airbnb_ny$price)

# Replace 'brookln' in borough by 'Brooklyn'
airbnb_ny$borough <- gsub("brookln", "Brooklyn", airbnb_ny$borough)

# Add a column for price range
airbnb_ny$price_range <- cut(as.integer(airbnb_ny$price), breaks= c(seq(0,1200, by= 100)))

# Remove 'NA' from price and price_range
airbnb_ny <- airbnb_ny %>% 
  filter(!is.na(price) & !is.na(price_range))

```

```{r, eval=FALSE}

# Remove unwanted columns
airbnb_ny <- airbnb %>% 
  select(id, host.id, host_identity_verified, neighbourhood.group, neighbourhood, lat, long, instant_bookable, cancellation_policy, room.type, Construction.year, price, review.rate.number)
  
# Rename the columns
airbnb_ny <- rename(airbnb_ny, listing_id = id, 
                    host_id = host.id,
                    host_verification_status = host_identity_verified,
                    borough = neighbourhood.group,
                    instant_booking = instant_bookable,
                    room_type = room.type,
                    construction_year = Construction.year, 
                    rating = review.rate.number)

# Remove entries with 'NA' from columns :lat, long, instant_booking, construction_year
airbnb_ny <- airbnb_ny %>% 
  filter(!is.na(lat) & !is.na(long) & !is.na(instant_booking) & !is.na(construction_year))

# Remove entries where host verification status is empty
airbnb_ny <- airbnb_ny %>% 
  filter(!host_verification_status == "")

# Remove entries where borough is empty
airbnb_ny <- airbnb_ny %>% 
  filter(!borough == "")

# Get the mean value of rating
airbnb_ny %>% 
  filter(!is.na(rating)) %>% 
  summarize(mean_rating = mean(rating))

# Replace NA in rating with mean value
airbnb_ny$rating[is.na(airbnb_ny$rating)] <- 3

# Get rid of $ and "," from price and cast as integer
airbnb_ny$price <- substr(airbnb_ny$price, 2, 6)
airbnb_ny$price <- gsub(",", "", airbnb_ny$price)
airbnb_ny$price <- as.integer(airbnb_ny$price)

# Replace 'brookln' in borough by 'Brooklyn'
airbnb_ny$borough <- gsub("brookln", "Brooklyn", airbnb_ny$borough)

# Add a column for price range
airbnb_ny$price_range <- cut(as.integer(airbnb_ny$price), breaks= c(seq(0,1200, by= 100)))

# Remove 'NA' from price and price_range
airbnb_ny <- airbnb_ny %>% 
  filter(!is.na(price) & !is.na(price_range))

```

### Data Visualization
```{r}
# Scatter plot: Rating vs price
ggplot(data = airbnb_ny, aes(x = rating, y = price))+
  geom_point()+
  facet_grid(~borough~room_type)+
  labs(x = "Rating",
       y = "Price",
       title = "Scatterplot: Rating vs price",
       subtitle = "Airbnb: Listings in New York")
```

```{r message=FALSE, warning=FALSE}
# Scatter plot: construction year vs price
ggplot(data = airbnb_ny, aes(x = construction_year, y = price ))+
  geom_point()+
  geom_smooth()+
  facet_grid(~borough~room_type)+
  labs(x = "Construction Year",
       y = "Price",
       title = "Scatterplot: Construction year vs price",
       subtitle = "Airbnb: Listings in New York")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

```{r}
# Number of listings per Borough
airbnb_ny %>% 
  group_by(borough) %>% 
  mutate(listings = n()) %>% 
  ggplot(aes(x= reorder(borough, -listings), label= borough, fill= borough ))+
  geom_bar(alpha= 0.7)+
  theme_bw()+
  geom_text(stat = 'count', position = position_stack(vjust = 0.9), aes(label=..count..), angle = 90)+
  labs(x = "Borough",
       y = "Number of listings",
       title = "Number of listings per Borough",
       subtitle = "Airbnb: Listings in New York")
```

```{r}
# Number of listings according to room type
airbnb_ny %>% 
  group_by(room_type) %>% 
  mutate(listings = n()) %>% 
  ggplot(aes(x= reorder(room_type, -listings), label = room_type, fill = room_type))+
  geom_bar(alpha= 0.7)+
  theme_bw()+
  geom_text(stat = 'count', position = position_stack(vjust = 0.9), aes(label=..count..), angle = 90)+
  labs(x = "Room type",
       y = "Number of listings",
       title = "Number of listings according to room type",
       subtitle = "Airbnb: Listings in New York")
```
```{r message=FALSE, warning=FALSE}
# Neighborhoods with highest average price for private room
# Unique neighborhoods
neighborhoods <- airbnb_ny %>% 
  select(neighbourhood, borough) %>% 
  group_by(neighbourhood, borough) %>% 
  summarise(count = n())

airbnb_ny %>% 
  group_by(neighbourhood) %>% 
  filter(room_type == "Private room") %>% 
  summarise(avg = as.integer(mean(price))) %>% 
  arrange(-avg) %>% 
  head(10) %>% 
  merge(y = neighborhoods, by = "neighbourhood", all=FALSE) %>% 
  ggplot(aes(x = reorder(neighbourhood, -avg), y =avg, label = avg, fill = borough))+
  theme_bw()+
  geom_bar(stat= 'identity', alpha=0.7)+
  geom_text(vjust= 1.5)+
  labs(x = "Neighbourhood",
       y = "Average price",
       title = "Neighborhoods with highest average price for private room",
       subtitle = "Airbnb: Listings in New York")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

```{r}
# Neighborhoods with highest average price for entire home/apt
airbnb_ny %>% 
  group_by(neighbourhood) %>% 
  filter(room_type == "Entire home/apt") %>% 
  summarise(avg = as.integer(mean(price))) %>% 
  arrange(-avg) %>% 
  head(10) %>% 
  merge(y = neighborhoods, by = "neighbourhood", all=FALSE) %>%  
  ggplot(aes(x = reorder(neighbourhood, -avg), y =avg, label = avg, fill = borough))+
  theme_bw()+
  geom_bar(stat= 'identity', alpha=0.7)+
  geom_text(vjust= 1.5)+
  labs(x = "Neighbourhood",
       y = "Average price",
       title = "Neighborhoods with highest average price for Entire home/apt",
       subtitle = "Airbnb: Listings in New York")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

```{r}
# Number of listings over price range for room type
ggplot(data = airbnb_ny, aes(x= price_range, label= price_range, fill= price_range))+
  geom_bar(alpha= 0.7)+
  theme_bw()+
  geom_text(stat = 'count', position = position_stack(vjust = 0.8), aes(label=..count..), angle = 90)+
  labs(x = "Price range",
       y = "Number of listings",
       title = "Number of listings over price range for room type",
       subtitle = "Airbnb: Listings in New York")+
  facet_wrap(~room_type)+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

```{r}
# Number of listings over a price range in borough
ggplot(data = airbnb_ny, aes(x= price_range, label= price_range, fill= price_range))+
  geom_bar(alpha= 0.7)+
  theme_bw()+
  geom_text(stat = 'count', position = position_stack(vjust = 0.7), aes(label=..count..), angle = 90)+
  labs(x = "Price range",
       y = "Number of listings",
       title = "Number of listings over a price range in borough",
       subtitle = "Airbnb: Listings in New York")+
  facet_wrap(~borough)+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

```{r message=FALSE, warning=FALSE}
# Average price by a room type in different boroughs
airbnb_ny %>% 
  group_by(room_type, borough) %>% 
  summarise(average_price= as.integer(mean(price))) %>% 
  ggplot(aes(x=borough, y= average_price, label= average_price, fill= borough, shape= room_type))+
  theme_bw()+
  geom_bar(stat= 'identity', alpha= 0.7, show.legend = FALSE)+
  facet_wrap(~room_type)+
  geom_text(vjust= 1.5)+
  labs(x = "Borough",
       y = "Average price",
       title = "Average price by a room type in different boroughs",
       subtitle = "Airbnb: Listings in New York")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```